Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-06-16 manuscript/manuscript.Rmd
fixes a few typos

### Model validation
To test each model's ability to forecast the population state we used out of sample predictions using leave-one-year-out cross validation. For both levels of modeling, we fit the vital rate models using observations from all years except one, and then used those fitted parameters in the population models to perform a one-step-ahead forecast for the year whose observations were withheld from model fitting. Within each observation year, several quadrats are sampled. So we made predictions for each observed quadrat in the focal year initialized with cover the previous year. Since we were making quadrat specific predictions we incorporated the group effect on the intercept for both models. We repeated this procedure for all 13 observation years, making 100 one-step-ahead forecasts for each quadrat-year combination with parameter uncertainty included via randomdrawd from the MCMC chain as described above. Random year effects were set to zero since year effects cannot be assigned to unobserved years.

This model validation allowed us to compare accuracy and precision of the two modeling approaches (individual-level versus population-level). We first calculated the median predicted cover across the 100 simulations for each quadrat-year and then calculated the absolute error as the difference between the observed cover for a given quadrat-year and the median prediction. To arrive at mean absolute error (MAE), we then averaged the absolute error within each species across the quadrat-year specific errors. We use MAE as our measure of accuracy. To measure precision calculated the distance between the upper and lower 90th quantiles of the 100 predictions as a measure of forecast precision and averaged this value over quadrat-years for each species.

### Testing sensitivity to climate covariates
Our main goal in this paper is to see if models based on aggregate level data are as sensitive to climate as models based on individual level data. So, with our fitted and validated models in hand, we ran simulations for each model type (IPM and QBM) under four climate perturbation scenarios: (1) observed climate, (2) precipitation increased by 1%, (3) temperature increased by 1%, and (4) precipitation and temperature increased by 1%. We ran this simulations for 2,500 time steps, enough to estimate equilibrium cover after discarding an initial 500 time steps as burn-in. Each simulation was run under two parameter scenarios: (1) using mean parameter estimates and (2) using randomly drawn parameters from the MCMC chain. We use (1) to detect the overall sensitivity of equilibrium cover to climate, and we use (2) to show the impact of model uncertainty on forecast precision.

@@

Results
-------
### Comparison of Forecast Models
The IPM had significantly lower overall error (MAE, mean absolute error) for two species (*H. comata* and *P. smithii*; Table 1). In no case did the QBM significantly outperform the IPM (Table 1). The IPM was consistently more precise than the QBM, with lower distances between the 90% quantiles across all species (Table 1). In general the IPM outperformed the QBM because it had (i) lower MAE for two species, (ii) statistically similar MAE for the other two species, and (iii) considerably more precise forecasts for all species.

### Forecasting Climate Change Impacts

>>>>

### Model validation
To test each model's ability to forecast the population state we made out of sample predictions using leave-one-year-out cross validation. For both levels of modeling, we fit the vital rate models using observations from all years except one, and then used those fitted parameters in the population models to perform a one-step-ahead forecast for the year whose observations were withheld from model fitting. Within each observation year, several quadrats are sampled. So we made predictions for each observed quadrat in the focal year initialized with cover the previous year. Since we were making quadrat specific predictions we incorporated the group effect on the intercept for both models. We repeated this procedure for all 13 observation years, making 100 one-step-ahead forecasts for each quadrat-year combination with parameter uncertainty included via randomdrawd from the MCMC chain as described above. Random year effects were set to zero since year effects cannot be assigned to unobserved years.

This model validation allowed us to compare accuracy and precision of the two modeling approaches (individual-level versus population-level). We first calculated the median predicted cover across the 100 simulations for each quadrat-year and then calculated the absolute error as the difference between the observed cover for a given quadrat-year and the median prediction. To arrive at mean absolute error (MAE), we then averaged the absolute error within each species across the quadrat-year specific errors. We use MAE as our measure of accuracy. To measure precision we calculated the distance between the upper and lower 90th quantiles of the 100 predictions and averaged this value over quadrat-years for each species.

### Testing sensitivity to climate covariates
Our main goal in this paper is to see if models based on aggregate level data are as sensitive to climate as models based on individual level data. So, with our fitted and validated models in hand, we ran simulations for each model type (IPM and QBM) under four climate perturbation scenarios: (1) observed climate, (2) precipitation increased by 1%, (3) temperature increased by 1%, and (4) precipitation and temperature increased by 1%. We ran the simulations for 2,500 time steps, enough to estimate equilibrium cover after discarding an initial 500 time steps as burn-in. Each simulation was run under two parameter scenarios: (1) using mean parameter estimates and (2) using randomly drawn parameters from the MCMC chain. We use (1) to detect the overall sensitivity of equilibrium cover to climate, and we use (2) to show the impact of model uncertainty on forecast precision.

@@

Results
-------
### Comparison of forecast models
The IPM had significantly lower overall error (MAE, mean absolute error) for three species (*B. gracilis*, *H. comata*, *P. smithii*; Table 1). In no case did the QBM significantly outperform the IPM (Table 1). The IPM was consistently more precise than the QBM, with lower distances between the 90% quantiles across all species (Table 1). In general the IPM outperformed the QBM because it had (1) lower MAE for three of the four species, (2) statistically similar MAE for the one other species, and (3) considerably more precise forecasts for all species.

### Sensitivity of models to climate

### 2015-06-16 manuscript/manuscript.tex
fixes a few typos

To test each model's ability to forecast the population state we used
out of sample predictions using leave-one-year-out cross validation. For
both levels of modeling, we fit the vital rate models using observations
from all years except one, and then used those fitted parameters in the
population models to perform a one-step-ahead forecast for the year
whose observations were withheld from model fitting. Within each
observation year, several quadrats are sampled. So we made predictions
for each observed quadrat in the focal year initialized with cover the
previous year. Since we were making quadrat specific predictions we
incorporated the group effect on the intercept for both models. We
repeated this procedure for all 13 observation years, making 100
one-step-ahead forecasts for each quadrat-year combination with
parameter uncertainty included via randomdrawd from the MCMC chain as
described above. Random year effects were set to zero since year effects
cannot be assigned to unobserved years.

This model validation allowed us to compare accuracy and precision of
the two modeling approaches (individual-level versus population-level).
We first calculated the median predicted cover across the 100
simulations for each quadrat-year and then calculated the absolute error
as the difference between the observed cover for a given quadrat-year
and the median prediction. To arrive at mean absolute error (MAE), we
then averaged the absolute error within each species across the
quadrat-year specific errors. We use MAE as our measure of accuracy. To
measure precision calculated the distance between the upper and lower
90th quantiles of the 100 predictions as a measure of forecast precision
and averaged this value over quadrat-years for each species.

@@

\subsubsection{Comparison of Forecast
Models}\label{comparison-of-forecast-models}

The IPM had significantly lower overall error (MAE, mean absolute error)
for two species (\emph{H. comata} and \emph{P. smithii}; Table 1). In no
case did the QBM significantly outperform the IPM (Table 1). The IPM was
consistently more precise than the QBM, with lower distances between the
90\% quantiles across all species (Table 1). In general the IPM
outperformed the QBM because it had (i) lower MAE for two species, (ii)
statistically similar MAE for the other two species, and (iii)
considerably more precise forecasts for all species.

\subsubsection{Forecasting Climate Change
Impacts}\label{forecasting-climate-change-impacts}

>>>>

To test each model's ability to forecast the population state we made
out of sample predictions using leave-one-year-out cross validation. For
both levels of modeling, we fit the vital rate models using observations
from all years except one, and then used those fitted parameters in the
population models to perform a one-step-ahead forecast for the year
whose observations were withheld from model fitting. Within each
observation year, several quadrats are sampled. So we made predictions
for each observed quadrat in the focal year initialized with cover the
previous year. Since we were making quadrat specific predictions we
incorporated the group effect on the intercept for both models. We
repeated this procedure for all 13 observation years, making 100
one-step-ahead forecasts for each quadrat-year combination with
parameter uncertainty included via randomdrawd from the MCMC chain as
described above. Random year effects were set to zero since year effects
cannot be assigned to unobserved years.

This model validation allowed us to compare accuracy and precision of
the two modeling approaches (individual-level versus population-level).
We first calculated the median predicted cover across the 100
simulations for each quadrat-year and then calculated the absolute error
as the difference between the observed cover for a given quadrat-year
and the median prediction. To arrive at mean absolute error (MAE), we
then averaged the absolute error within each species across the
quadrat-year specific errors. We use MAE as our measure of accuracy. To
measure precision we calculated the distance between the upper and lower
90th quantiles of the 100 predictions and averaged this value over
quadrat-years for each species.

@@

\subsubsection{Comparison of forecast
models}\label{comparison-of-forecast-models}

The IPM had significantly lower overall error (MAE, mean absolute error)
for three species (\emph{B. gracilis}, \emph{H. comata}, \emph{P.
smithii}; Table 1). In no case did the QBM significantly outperform the
IPM (Table 1). The IPM was consistently more precise than the QBM, with
lower distances between the 90\% quantiles across all species (Table 1).
In general the IPM outperformed the QBM because it had (1) lower MAE for
three of the four species, (2) statistically similar MAE for the one
other species, and (3) considerably more precise forecasts for all
species.

\subsubsection{Sensitivity of models to
climate}\label{sensitivity-of-models-to-climate}
